### 2015-02-03 d03-acronyms.tex

%-----------------------------------------
% acronyms
%-----------------------------------------

\newacronym{psha}{PSHA}{Probabilistic Seismic Hazard Analysis}
\newacronym{dsha}{DSHA}{Deterministic Seismic Hazard Analysis}
\newacronym{GMPE}{\gls{gmpe}}{\glsdesc{gmpe}}
\newacronym{mfd}{MFD}{Magnitude Frequency Distribution}
\newacronym{gr}{GR}{Gutenberg-Richter}
\newacronym{cnn}{CNN}{Coupled-Nearest-Neighbour}
\newacronym{va}{r.v.}{random variable}
\newacronym{iid}{i.i.d.}{independent and identicaly distributed}
\newacronym{pdf}{density}{probability density function}
\newacronym{pmf}{pmf}{probability mass function}
\newacronym{kde}{KDE}{Kernel Density Estimation}

\newacronym{pga}{PGA}{peak ground-motion acceleration}

\newacronym{isc}{ISC}{International Seismological Centre}
\newacronym{gem}{GEM}{Global Earthquake Model}
\newacronym{opensha}{openSHA}{Open Source Seismic Hazard Analysis}
\newacronym{oq}{OQ}{Openquake}
\newacronym{eeri}{EERI}{Earthquake Engineering Research Institute}
\newacronym{cgmw}{CGMW}{Comission for Geological Map of the World}

\newacronym{hl}{HazardLib}{hazard library}
\newacronym{rl}{RiskLib}{risk library}
\newacronym{nrml}{NRML}{Natural-hazard and Risk Markup Language}
\newacronym{hmtk}{HMTK}{Hazard Modeller's Toolkit}
\newacronym{oqp}{oq-platform}{Plataforma web de Interação com o OQ}
\newacronym{oqe}{oq-engine}{OQ hazard and risk calculator}

\newacronym{obsis}{ObSis}{Seismological Observatory}
\newacronym{unb}{UnB}{Brasilia University}
\newacronym{iag}{IAG}{Institute of Astronomy, Geophysics and Atmosferical Sciences}
\newacronym{usp}{USP}{São Paulo University}
\newacronym{ipt}{IPT}{Institute of Technological Research}
\newacronym{unesp}{UNESP}{São Paulo State University}
\newacronym{ufrn}{UFRN}{Federal University of Rio Grande do Norte}

\newacronym{bsb}{BSB}{Brazilian Seismic Bulletin}
\newacronym{csv}{CSV}{Common separated values}

\newacronym{bsb2013}{{BSB-2013.08}}{Brazilian Seismic Bulletin version 2013.08}
\newacronym{bsb2014}{{BSB-2014.11}}{Brazilian Seismic Bulletin version 2011.11}
\newacronym{iscgem}{{\gls*{isc}-\gls*{gem}}}{ISC-GEM catalog for South America}