### 2015-02-10 manuscript.tex
completeness figure and map + clean abstract and review intro

%!TEX program = pdflatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% American Geophysical Union (AGU)
% LaTeX Template
% Version 1.0 (3/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% The AGUTeX class and agu-ps referencing style were created and are owned
% by AGU: http://publications.agu.org/author-resource-center/author-guide/latex-formatting-toolkit/
%
% This template has been modified from the blank AGU template to include
% examples of how to insert content and drastically change commenting. The
% structural integrity is maintained as in the original blank template.
%
% Important notes:
% This template retains extensive commenting from the AGU template. It is heavily
% advised you read these comments and follow them in order to insure a speedy
% submission process.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified March 2013
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Step 1: Set the \documentclass

% There are two options for article format: two column (default) and draft.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.

% Choose the journal abbreviation for the journal you are submitting to:

% jgrga	JOURNAL OF GEOPHYSICAL RESEARCH
% gbc	GLOBAL BIOCHEMICAL CYCLES
% grl	GEOPHYSICAL RESEARCH LETTERS
% pal	PALEOCEANOGRAPHY
% ras	RADIO SCIENCE
% rog	REVIEWS OF GEOPHYSICS
% tec	TECTONICS
% wrr	WATER RESOURCES RESEARCH
% gc	GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw	SPACE WEATHER
% ms	JAMES
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

% \documentclass[draft, grl]{agutex}
\documentclass[grl]{agutex}

% To create numbered lines:

% If you don't already have lineno.sty, you can download it from http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/ (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN). Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno} and \linenumbers*[1]command, below:

% \usepackage{lineno}
% \linenumbers*[1]

%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables

% DO NOT USE \psfrag or \subfigure commands.

%  Figures and tables should be placed AT THE END OF THE ARTICLE, after the references.

%  Uncomment the following command to include .eps files (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
\usepackage{graphicx}

% Substitute one of the following for [dvips] above if you are using a different driver program and want to proof your illustrations on your machine:
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]

%  Uncomment the following command to allow illustrations to print when using Draft:
\setkeys{Gin}{draft=false}

% See how to enter figures and tables at the end of the article, after references.
\usepackage[utf8]{inputenc}
\usepackage{float}
%\usepackage{caption}
%\usepackage{subcaption}

% \usepackage[round,sort,nonamebreak]{natbib} % citação bibliográfica textual(plainnat-ime.bst)
% \bibpunct{(}{)}{;}{a}{\hspace{-0.7ex},}{,} % estilo de citação. Veja alguns exemplos em http://merkel.zoneo.net/Latex/natbib.php

\usepackage{amsfonts}
\usepackage{textcomp}

%-----------------------------------------------------------------------------
%	colored references
%-----------------------------------------------------------------------------
% \usepackage[usenames,svgnames,dvipsnames]{xcolor}
\usepackage[a4paper,top=2.5cm,bottom=3.0cm,left=2.0cm,right=2.0cm]{geometry}
% \usepackage[colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=Green,bookmarksopen=true]{hyperref}
% \usepackage[all]{hypcap}      % soluciona o problema com o hyperref
%-----------------------------------------------------------------------------
%	Glossaries
%-----------------------------------------------------------------------------

% Glossaries
%\usepackage[sanitize=none,acronym,toc]{glossaries}
\usepackage[acronym,toc]{glossaries}
%\usepackage[acronym,toc]{glossaries}
% Define a new glossary type
\newglossary[slg,toc]{symbols}{sym}{sbl}{Symbols List}
\newglossary[elg]{equations}{eqn}{eql}{Equations}
\makeglossaries

% ---------------------------------------------------------------------------- %
% Dicionario de Termos:
\input{d01-terms}
\input{d02-symbols}
\input{d03-acronyms}
\input{d04-equations}
% ---------------------------------------------------------------------------- %

% ---------------------------------------------------------------------------- %
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\erf}{erf}
% ---------------------------------------------------------------------------- %

%----------------------------------------------------------------------------------------
%	RUNNING HEAD AND CORRESPONDING AUTHOR
%----------------------------------------------------------------------------------------

% Author names in capital letters:
\authorrunninghead{PIRCHINER M. and ASSUMP\c{C}\~AO M.}

%------------------------------------------------

% Shorter version of title entered in capital letters:
\titlerunninghead{Smoothed seismicity Brazilian PSHA}

%------------------------------------------------

% Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: 
Marlon Pirchiner, 
Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
(marlon.pirchiner@iag.usp.br)}

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\title{Smoothing seismicity techniques applied to
seismic source characterization and
probabilistic seismic hazard analysis in Brazil}

%----------------------------------------------------------------------------------------
%	AUTHORS AND AFFILIATIONS
%----------------------------------------------------------------------------------------

% Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote; matching \altaffiltext will appear at bottom of page.

\authors{Marlon Pirchiner\altaffilmark{1,2} 
		and Marcelo Assumpção\altaffilmark{1}
		}

\altaffiltext{1}{Department of Geophysics,
Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG),
University of São Paulo (USP), São Paulo, Brazil.}

\altaffiltext{2}{Applied Math School (EMAp), Getúlio Vargas Foundation (FGV), Rio de Janeiro, Brazil.}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% Do NOT include any \begin...\end commands within the body of the abstract.

\begin{abstract}
Intraplate seismicity is a hard problem to solve in tectonics, mainly during the lack of observed seismicity in many areas where seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.
% Intraplate seismicity is a hard problem to solve in tectonics. In general, the earthquake occurrence is lower than others ``active'' regions. In the Brazilian case, there is a lack of observed seismicity due the short (in geological scale) observation timeframe. In addition, the populational density decreases towards to interior of the country and changes the reporting capabilities historicaly.

% This diffused intraplate seismicity pattern is often modeled as polygons representing different seismic zones or area sources for \gls{psha} purposes. Their geometries are defined by expert opinions.

% Other possible approach to deal this is develop zoneless or smoothed  models for the seismicity, conside

Three smoothing seismicity techniques are reviewed and their application to the Brazilian seismicity and a seismic hazard assessment are discussed.

All these smoothing methods are based on a \gls{kde}. The main difference between them is the bandwidth selection process.

The method of \citet{frankel_1995} uses a \emph{fixed} bandwidth in space. \citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}. \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event.

This last one was proposed initially to take care of the background seismicity rate in the context of long-term earthquake forecast which time-independence assumption able us to use this recurrence rate to define a grid of seismic point sources as the input source model to compute the hazard.

The results shown that all methods clearly define some known seismic regions and could be used in the Brazilian seismicity. All methods improve the spatial resolution of the GSHAP model, the only previously available for the region.

% The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.

All the hazard computations was performed on OpenQuake \citep{pagani_2014} open source software also from GEM Foundation.

\end{abstract}

%------------------------------------------------------------------------------
%	ARTICLE CONTENT
%------------------------------------------------------------------------------
% The body of the article must start with a \begin{article} command
% \end{article} must follow the references section, before the figures and tables.

\begin{article}

\section{Introduction}

\subsection{Intraplate seismicity}

COMENTAR CAPITULO DO LIVRO

\subsection{Previous studies in Brazil}

Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984, berrocal_1996, assumpcao_et_al_2014}, Brazil has no seismic hazard map at national scale.

The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region. However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GSHAP-based building codes.

A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better empirical and updated definition of the seismicity and hazard levels.

% definition of seismic zones. For this reason an updated hazard map is needed.

\subsection{Smoothed seismicity}

Although the standard \citet{cornell_1968} and \citet{mcguire_1976} methodology often used to assess seismic hazard is using expert opinions to pre-defined seismic zones, the goal of this paper is to review a few smoothing seismicity methods based on \gls{kde} which results in a grid of smoothed seismicity without define any seismic zone polygon.

The seismicity \gls{kde} or the smoothed seismicity grid allow to define a grid of seismic point sources. The seismic rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods, but the other parameters, like \emph{b}-values, minimum and maximum magnitudes, rupture and depth distributions, must be assigned in other ways.

To perform the \gls{kde}, both its shape and bandwidth needs to be defined. Most kernels are Gaussian or power-laws or one of their variants. The main distinction between the methods described here is the bandwidth selection process.

\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} smoothing technique with a \emph{fixed} smoothing-distance bandwidth to estimate the seismicity rate smoothing the 2D histogram counting. This version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1}. There is also a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which sum the contribution of each earthquake kernel on given cell. \citet{woo_1996} proposes a \emph{magnitude-dependent} kernel bandwidth based on the nearest neighbor mean distance in magnitude bins. \citet{helmstetter_2012} proposes a \emph{locally-adaptive} bandwidth, based on kernel estimations for the background seismicity rate in long-term forecasts on both space and time dimensions.

Using the background seismic rate computed to the forecast and assuming time-independence of that, the same assumption of PSHA, it is possible, preserving their differences \citep{marzocchi_2011}, to use these forecasted seismicity values on seismic sources characterization \citep{weatherill_pagani_2014}.

\subsection{Hazard computation}

To perform the hazard computation, the OpenQuake suite \citep{pagani_2014} was chosen. The modelling tool was the \gls{hmtk} also open available by \gls{gem} scientific board. The Frankel method was already implemented on the toolkit and the other two was implemented in the context of this paper and they are free available on a public repository.

The earthquake catalog data comes from the Brazilian seismological research authorities.

\subsection{Specific purposes}

The specific purposes of this paper are: (i) review 3 distincts smoothing/zoneless methods to describe spatially the seismicity rate, (ii) update the seismic hazard studies at national scale by proposing a new hazard model,
(iii) use the GEM-OpenQuake tools to perform hazard calculation and (iv) add and share with the community two more new smoothing methods into the \gls{hmtk}.

To aim this, the Brazilian seismicity data is presented including some overview plots. Next the method's theory is presented follow by the specific decisions and optimizations performed on each method modelling. And last, the results, conclusions and further considerations are discussed.

%------------------------------------------------
%
% seismicity data
%
%------------------------------------------------

\section{Brazilian seismicity data}

Data on intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo \citep{bsb_2014}. This catalog is a joint compilation from different sources.

\subsection{Hypocenters}

For events up to 1981, the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with participation of Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT). From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica. Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP). Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.

Today the catalog is distributed in two ways. One, which could be called \emph{raw}, contains all compiled information, such as: events outside the country which were felt in Brazil, errors in previous versions and also non-seismic events (known quarry blasts, sonic boom, etc). The other way could be called \emph{clean}, where all non-seismic and low quality events were removed. This last one was used in this paper and the total number of entries is [CALCULAR O NUMERO DE EVENTOS NO CATALOGO].

This catalog only cover crustal events. Andean subduction earthquakes deeper than 50km in the Brazil-Peru border was discarded. The epicenters of the catalog are shown in figure \ref{fig_seismicity}.

\subsection{Magnitudes}

Most magnitudes were computed as $mb$ type or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude. For historical events magnitudes were estimated from macroseismic data (felt area and maximum intensity) \citep{berrocal_1984}. Since most \gls{gmpe} are based on $M_W$ moment scale magnitude, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}. REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS. (Johnston)
% Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst almost magnitude values on \citet{bsb_2014} catalog have magnitude values lower than that.

\subsection{Catalog checking}

Figure \ref{fig_overview} shows the distribution of depths, day of the week and origin hour following \citet{gulia_2012}. Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10km.

Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day of the week, and a slightly trend to record earthquakes during the night. This could suggest some influence of daily noise level.

The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara \citep{takeya_1989} and Porto dos Gaúchos \citep{barros_2009}. The increasing number of records after 1960 is due to the operation of global, regional and local stations.

\subsection{Declustering procedures}

To preserve the Poissonian assumption about the independence of the events, several declustering algorithms were tried and the results as shown in figure \ref{fig_decluster}a. The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uenthal \citep{van_stiphout_2012}.
In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window.

In the case of Brazilian catalog the methods did not present large differences at the final number of earthquakes (fig. \ref{fig_decluster}a) and the Gardner-Knopoff/Uhrhammer combination was chosen to maximize the number of events.

The map of figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where long earthquake series tend to occur more frequently.

\subsection{Frequency-magnitude distribution}

Figure \ref{fig_occurrence} presents the catalog \gls{mfd}.
A general $b=1$ fit is also plotted, and the completeness magnitude on the incremental distribution is about 3.0.

\subsection{Catalog completeness}

The completeness of Brazilian seismicity data has evident time and space dependence.

Brazil just recently was installed its permanent seismic network \citep{pirchiner_2011}. The number and quality of stations increased over time. The population density, previously concentrated along the coast increased in other regions changing the historical detectability.

There are many methodologies to assessment of the temporal and spatial completeness \citep{stepp_1972, mignan_woessner_2012, ogata_1993, wiemer_wyss_2000, cao_gao_2002, stucchi_2004, woessner_2005, mignan_2011, mignan_2012, vorobieva_2013, mignan_2013, nasir_2013, mignan_2014}, or explaining how to compute the real \gls{pde} from the waveforms \citep{schorlemmer_2008}, but the time and the effort to proceed this exhaustive study of the \glsdesc{sym:Mc}, despite its relevance, was not intended to be done on the scope of this paper.

The waveforms from the earliest years was not in digital form. The histories of most regional (temporary) stations are not easy available. In addition, the low number of earthquakes in the catalog also difficults some high resolution statistical methods. All of this hinders the direct application of most that methodologies.

For this reason, the spatial completeness was not considered and the temporal completeness parameters for the whole country was defined by an easy cumulative rate criteria using historical information about the network evolution and the \citet{stepp_1972} plot (fig. \ref{fig_completeness}a). TALK ABOUT BRASILIA STATION???

Figure \ref{fig_completeness}b shows the magnitude time distribution in a scatter plot as the cumulative number of events which is important to perform the \gls{tcef} \citep{nasir_2013}. The linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend. Figure \ref{fig_completeness}b also show three (from many) possible completeness tables to allow quick check their distinctions. We chosen the ``simplest'' one.

IMPROVE THIS DECISION EXPLANATION

%------------------------------------------------
%
% smooothing methods
%
%------------------------------------------------
\section{Smoothing methods}
Three smoothing methods are now briefly explained.

\subsection{Frankel: fixed bandwidth}

\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a ``correlation distance'' $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$. $n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.

\subsection{Woo: magnitude-dependent bandwidth}

\citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, located at $\boldsymbol{r}_i$, to the seismicity rate $R$ (number of earthquakes per year and unity area) on the cell centered in $\boldsymbol{r}$, depending on the magnitude $m$:
\begin{equation}
	\ensuremath{
		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
											{T({\gls{sym:ri})}},
	}
	\label{eq:Rrm}
\end{equation}
where $N$ is the number of earthquakes $i$ in the magnitude bin $m \pm \mathrm{d}m$,
and $T(\gls{sym:ri})$ is the completeness time of magnitude $m$ observed on \gls{sym:ri}.

Any kernel could be applied on that definition. I used used the \citet{kagan_knopoff_1980} kernel for infinite spatial domain as described by \citet{woo_1996}:
\begin{equation}
	\ensuremath{
		K_{KK}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
							\left[ 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right]^{-\gls{sym:aW}},
	}
	\label{eq:k_kj}
\end{equation}
where \gls{sym:aW} is \glsdesc{sym:aW} \citep{verejones_1992}.

VERIFICAR SE É O CASO DE APRESENTAR OUTRAS FORMAS DE KERNEL. DAR PREFERENCIAS AOS REALMENTE UTILIZADOS.

To compute the magnitude-dependent bandwidth function \gls{sym:hm}, \citet{woo_1996} suggested the follow relation
\begin{equation}
	\ensuremath{
		h(m\arrowvert \gls{sym:a0}, \gls{sym:a1}) = \gls{sym:a0}e^{\gls{sym:a1}m},
	}
	\label{eq:hm}
\end{equation}
where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$, as illustrated on figure \ref{fig_woo_bandwidth}.

Just to compare, in figure \ref{fig_woo_bandwidth} it is shown the functions computed by \citet{beauval_2003} for Norway and Spain. The angular coefficient ($a_1$) is in the same range of the others but the general mean distance between events into the same magnitude bin is higher than other countries. It is in some way expected. Brazil is quite large than mentioned countries.

\subsection{Helmstetter and Werner: locally-adaptive space and time bandwidths}

Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to characterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.

\citet{helmstetter_2012} proposes a seismicity model space and time dependent using kernels for space and time independently:
\begin{equation}
	\ensuremath{\gls{sym:R} = \sum_{i=1}^{N}{ \frac{1}{h_i\,{d_i}^2} \gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms01}
\end{equation}
where \gls{sym:R} is the \glsdesc{sym:R},
	  $K_t$ is the \glsdesc{sym:Kt},
	  $K_r$ is the \glsdesc{sym:Kr}.

As they are interested in forecast, just past time $t_i < t$ need to be considered. Also the observation completeness is taken in account by a set of weights $w$ in this follow way:
\begin{equation}
\ensuremath{\gls{sym:R} = \gls{sym:Rmin} + \sum_{t_i < t}{
	\frac{2\,w(\boldsymbol{r}_i,t_i)}{h_i\,{d_i}^2}
			\gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms02}
\end{equation}
where \gls{sym:Rmin} is the \glsdesc{sym:Rmin}, positive, allowing earthquakes to occur where their never occurred yet.

The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection. This is the expression used to compute it:
\begin{equation}
	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
	\right] } },
	\label{eq:helms_wi}
\end{equation}
where \gls{sym:wi} is the \glsdesc{sym:wi} factor computed on location $\boldsymbol{r}$ on instant $t$,
	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
	  \gls{sym:Mc_rt} is the \glsdesc{sym:Mc_rt},
	  \gls{sym:Md} is the \glsdesc{sym:Md}.

These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$. They also could easily take into account the space and time completenes and \emph{b}-value fluctuations.

\subsubsection{Local bandwidth computation}

The method implemented to compute the space and time bandwidths for each earthquake was the \gls{cnn} \citep{helmstetter_2012} expressed as:
\begin{equation}
	\ensuremath{
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
	 \gls{sym:dk} is the \glsdesc{sym:dk} and
	 \gls{sym:hk} is the \glsdesc{sym:hk}.

The bandwidths are defined locally by this simple optimization process. It could be small on high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.

\subsubsection{Stationary seismic rate}
After compute the model parameters it is completely defined, and the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ can be computed on the way proposed by \citet{helmstetter_2012} taking the median value for some location $\boldsymbol{r}_0$ over all considered time window:
\begin{equation}
	\ensuremath{
		\bar{R}(\boldsymbol{r}_0) = \text{Median}\left[R(\boldsymbol{r}_0, t)\right].
	}
	\label{eq:stationary_rate}
\end{equation}

The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures. This is one of the most important achievement of this methodology.

\subsubsection{Maximum likelihood optimization}

The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}. To optimize these parameters the catalog is divided in two parts: one for \emph{learning} about the parameters and other to test them. The \emph{testing} catalog will able us to check the model prediction performance derived from its chosen parameters. The best model parameters will maximize the prediction model capacity likely the target catalog.

If the earthquake occurrence could be modeled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
\begin{equation}
	\ensuremath{
		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
	}
	\label{eq:poisson_probability}
\end{equation}

Then, over all cells, the log-likelihood, to be maximized, between model prediction (from the learning catalog) and the observed earthquakes (on the testing catalog) is written as
\begin{equation}
	\ensuremath{
		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
	}
	\label{eq:loglik}
\end{equation}
where \gls{sym:Np} is the \glsdesc{sym:Np} and
	  \gls{sym:nxy} is the \glsdesc{sym:nxy}.

The model parameters $R_{min}$, $a_{cnn}$ and $k_{cnn}$ should be optimized by the maximization of log-likelihood $L$.

%------------------------------------------------

\section{Results}

.

\subsection{Simulations}

\subsubsection{Simulation 1}

.

%------------------------------------------------

\section{Discussion}

.

%----------------------------------------------------------------------------------------
%	APPENDICES (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here

% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing  \appendix

% \section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title

\appendix

\section{Appendix Title}

.

%----------------------------------------------------------------------------------------
%	GLOSSARY OR NOTATION (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%------------------------------------------------------------------------------
\begin{acknowledgments}
The author thanks specially to 
%Marcelo Assumpção for the support and critical discussion as well 
all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
\end{acknowledgments}

%------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%------------------------------------------------------------------------------

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}

% Or,

% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.

% Follow these steps:
% 1. Run LaTeX on your LaTeX file.

% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.

% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.

% 4. Comment out the old \bibliographystyle and \bibliography commands.

% 5. Run LaTeX on your new file before submitting.

% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

% ---------------------------------------------------------------------------- % Bibliografy
% ----------------------------------------------------------------------------
\bibliographystyle{BibTeX/agufull08} 	% citação bibliográfica textual
\bibliography{bib/bibliografia}  		% associado ao arquivo:

% \begin{thebibliography}{}

% \providecommand{\natexlab}[1]{#1}
% \expandafter\ifx\csname urlstyle\endcsname\relax
%   \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
%   \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
%   \urlstyle{rm}\Url}\fi

% \bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
% Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
%   logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
%   Comp.}, \textit{56}(193), 119--139.

% \bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
% Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
%   Scattering Theory}, John Wiley, New York.

% \bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
%   Wendland}}]{StephanHsiao}
% Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
%   problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
%   Appl. Math.}, \textit{34}(1), 1--19.

% \bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
% Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
%   components and line integrals of currents in modified edge representation,
%   \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

% \end{thebibliography}

% Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

\end{article}

%------------------------------------------------------------------------------
%	FIGURES AND TABLES
%------------------------------------------------------------------------------

%% Enter Figures and Tables here:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information should be placed in footnotes below the table.
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}
% \noindent\includegraphics[width=20pc]{samplefigure.eps}
% \caption{Caption text here}
% \label{figure_label}
% \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% \begin{figure}
% 	\begin{center}
% 	\begin{table}[H]
% 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
% 		\footnotesize(a)
% 		&
% 		\footnotesize(b)
% 		\\
% 		\includegraphics[width=0.45\textwidth]{placeholder} &
% 		\includegraphics[width=0.45\textwidth]{placeholder}
% 		\end{tabular}
% 	\end{table}
% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
% 	\label{placeholder}
% 	\end{center}
% \end{figure}

% \begin{figure}
% 	\begin{center}
% 	\begin{table}[H]
% 		\begin{tabular}{ p{0.33\textwidth}
% 						 p{0.33\textwidth}
% 						 p{0.33\textwidth} }
% 		\footnotesize(a) &
% 		\footnotesize(b) &
% 		\footnotesize(c)
% 		\\
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
% 		\end{tabular}
% 	\end{table}
% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
% 	\label{placeholder2}
% 	\end{center}
% \end{figure}

% See below for how to make sideways figures or tables.

\newpage

% =========
% intro
% =========

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_seismicity}
	\caption{Brazilian seismicity from 1767 to 2014. Source: \citep{bsb_2014}.}
	\label{fig_seismicity}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.20\textwidth}
						 p{0.40\textwidth}
						 p{0.40\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
		\end{tabular}
	\end{table}
	\caption{Catalogue overview. The histogram (a) shows the depth distribution. The histograms (b) and (c) represent the distributions of weekday and hour which earthquakes occur respectly. Dashed lines represent the mean value.}
	\label{fig_overview}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.99\textwidth]{z_img_hmtk_bsb2014_11_rate}
		\\
		\footnotesize(b)
		& \includegraphics[width=0.90\textwidth]{z_img_time_mag_count_br_1960}
		\end{tabular}
	\end{table}
	\caption{Earthquakes by year (a) and discriminated by magnitude (b).}
	\label{fig_records}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.70\textwidth]{z_img_hmtk_bsb2014_11_decluster}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.60\textwidth]{z_img_hmtk_bsb2014_11_clusters}
		\end{tabular}
	\end{table}
	\caption{Decluster evaluation (a) using distinct algorithms and windows. There is no significant differences between final results. The map (b) of clusters from selected method shows places where earthquake sequences tends to occur more frequently.}
	\label{fig_decluster}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
        & \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_stepp}
		\\
		\footnotesize(b)
		& \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_scatter}
		\end{tabular}
	\end{table}
	\caption{Completeness evaluation. 
	The Stepp plot (a) was made using $\Delta_{mag}=0.5$ and $\Delta_{time}=2$ as magnitude and time bins. The magnitude-time scatter plot (b) shows in addition, three diferent temporal completess models and some cumulative rates to better perform a visual analysis.}
	\label{fig_completeness}
	\end{center}
\end{figure}

\begin{figure}
    \includegraphics[width=.99\linewidth]{z_img_hmtk_bsb2014_11_recurrence}
    \caption{Annual earthquake recurrence.}
    \label{fig_occurrence}
\end{figure}

% =========
% Methods
% =========

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_woo_bandwidth}
	\caption{Magnitude dependence bandwidth.}
	\label{fig_woo_bandwidth}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_woo_kernels}
	\caption{\citet{kagan_knopoff_1980} kernel shapes for some magnitude values.}
	\label{fig_woo_kernels}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_hidi}
	\caption{Local bandwidth example.}
	\label{fig_helmstetter_hidi}
\end{figure}

% TODO add HI_DI histograms

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.99\textwidth]{z_img_hi_histogram}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.99\textwidth]{z_img_di_histogram}
		\end{tabular}
	\end{table}
	\caption{(a) $h_i$ and (b) $d_i$ distributions from \gls{bsb}.}
	\label{fig_hidi_histogram}
	\end{center}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_stationary_a}
	\caption{Stationary seismic rate}
	\label{fig_helmstetter_stationary_rate}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_catalogues}
	\caption{Learning and testing catalogs.}
	\label{fig_helmstetter_catalogues}
\end{figure}

% =========
% Smoothed Rates
% =========

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.33\textwidth}
						 p{0.34\textwidth}
						 p{0.33\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
		\includegraphics[height=5cm]{z_img_a_woo} &
		\includegraphics[height=5cm]{z_img_a_helmstetter}
		\end{tabular}
	\end{table}
	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
	\label{fig_rates}
	\end{center}
\end{figure}

% =========
% Results
% =========

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.33\textwidth}
						 p{0.34\textwidth}
						 p{0.33\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_pga_frankel}  &
		\includegraphics[height=5cm]{z_img_pga_woo} &
		\includegraphics[height=5cm]{z_img_pga_helmstetter}
		\end{tabular}
	\end{table}
	\caption{PGA (poe 10\%/50y) comparission for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
	\label{fig_pga}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\footnotesize(a)
		&
		\footnotesize(b)
		\\
		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
		\end{tabular}
	\end{table}
	\caption{Previous available Brazillian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
	\label{placeholder}
	\end{center}
\end{figure}

% TABLES

\begin{table}[H]
	\caption{BSB 2014.11 $M_W$ proxies \emph{ad-hoc} guidelines.}
	\begin{tabular}{l l l}
		\hline
		\textbf{magnitude source} & \textbf{rule} & \textbf{uncertainty}\\
		\hline
		% mb or equival & $$ M_W = 1.21m - 0.76 $$  & 0.3 \\
		% 		% magnitude from $A_f$  & $$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 8.4\times10^{-4}A_f^{1/2}& 0.4 \\
		$mb$ or $mR$
		&$ M_W(m) = 1.121 m - 0.76 $
		& 0.3 \\
		Area felt $A_f$
		&$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 0.00084\sqrt{A_f}$& 0.4 \\
		Maximum intensity $I_0$
		& $mb(I_0) = 1.21 + 0.45 I_0$ then $M_W(m)$
		& 0.6 \\
		$mb$ and $A_f$
		& $M_W(m, A_f) = 0.7 M_W(m) + 0.3 M_W(A_f)$
		& 0.33 \\
		\hline
	\end{tabular}
	\label{tab:guidelines}
\end{table}

\begin{table}[H]
	\caption{Temporal completeness comparision. For each magnitude interval,
	the completess was computed by the \citet{stepp_1972} method and the respective assigned value.}
	\begin{tabular}{c c c}
		\hline
	\textbf{magnitude} & \textbf{year (Stepp)} & \textbf{year (simplest)} \\
		\hline
		2.0 & 1982 	& 1885 \\
		2.5 & 1980 	& 1965 \\
		3.0 & 1976 	&  \\
		3.5 & 1969 	& 1960 \\
		4.0 & 1955 	&  \\
		4.5 &  		& 1935 \\
		5.0 & 1952 	&  \\
		5.5 & 1927 	&  \\
		6.0 &  		&  \\
		6.5 &  		&  \\
		\hline
	\end{tabular}
	\label{tab:completeness}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 4 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2 \]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with the first character following the equal sign on the first line.

% Use an \hspace{} command to insert horizontal space into your equation if necessary. Place an appropriate unit of measure between the curly braces, e.g. \hspace{1in}; you may have to experiment to achieve the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}