### 2015-05-17 manuscript.tex
bibliography, more comments in the text using the figures and improoving legends (section 2)

%!TEX program = pdflatex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% American Geophysical Union (AGU)
% LaTeX Template
% Version 1.0 (3/6/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% The AGUTeX class and agu-ps referencing style were created and are owned
% by AGU: http://publications.agu.org/author-resource-center/author-guide/latex-formatting-toolkit/
%
% This template has been modified from the blank AGU template to include
% examples of how to insert content and drastically change commenting. The
% structural integrity is maintained as in the original blank template.
%
% Important notes:
% This template retains extensive commenting from the AGU template. It is heavily
% advised you read these comments and follow them in order to insure a speedy
% submission process.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified March 2013
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Step 1: Set the \documentclass

% There are two options for article format: two column (default) and draft.

% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.

% Choose the journal abbreviation for the journal you are submitting to:

% jgrga	JOURNAL OF GEOPHYSICAL RESEARCH
% gbc	GLOBAL BIOCHEMICAL CYCLES
% grl	GEOPHYSICAL RESEARCH LETTERS
% pal	PALEOCEANOGRAPHY
% ras	RADIO SCIENCE
% rog	REVIEWS OF GEOPHYSICS
% tec	TECTONICS
% wrr	WATER RESOURCES RESEARCH
% gc	GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw	SPACE WEATHER
% ms	JAMES
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

\documentclass[draft, grl]{agutex}
% \documentclass[grl]{agutex}

% package used by \citep and \citet
% \usepackage[sort&compress,square,comma,authoryear]{natbib}

% makes color citations
% \usepackage[dvips,dvipdfm,colorlinks=true,urlcolor=blue,citecolor=red,linkcolor=red,bookmarks=true]{hyperref}

% To create numbered lines:

% If you don't already have lineno.sty, you can download it from http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/ (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN). Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno} and \linenumbers*[1]command, below:

% \usepackage{lineno}
% \linenumbers*[1]

%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables

% DO NOT USE \psfrag or \subfigure commands.

%  Figures and tables should be placed AT THE END OF THE ARTICLE, after the references.

%  Uncomment the following command to include .eps files (comment out this line for draft format):
%\usepackage[dvips]{graphicx}
\usepackage{graphicx}

% Substitute one of the following for [dvips] above if you are using a different driver program and want to proof your illustrations on your machine:
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]

%  Uncomment the following command to allow illustrations to print when using Draft:
\setkeys{Gin}{draft=false}

% See how to enter figures and tables at the end of the article, after references.
\usepackage[utf8]{inputenc}
\usepackage{float}
%\usepackage{caption}
%\usepackage{subcaption}

% \usepackage[round,sort,nonamebreak]{natbib} % citação bibliográfica textual(plainnat-ime.bst)
% \bibpunct{(}{)}{;}{a}{\hspace{-0.7ex},}{,} % estilo de citação. Veja alguns exemplos em http://merkel.zoneo.net/Latex/natbib.php

\usepackage{amsfonts}
\usepackage{textcomp}

%-----------------------------------------------------------------------------
%	colored references
%-----------------------------------------------------------------------------
% \usepackage[usenames,svgnames,dvipsnames]{xcolor}
\usepackage[a4paper,top=2.5cm,bottom=3.0cm,left=2.0cm,right=2.0cm]{geometry}
% \usepackage[colorlinks=true,citecolor=DarkGreen,linkcolor=NavyBlue,urlcolor=DarkRed,filecolor=Green,bookmarksopen=true]{hyperref}
% \usepackage[all]{hypcap}      % soluciona o problema com o hyperref
%-----------------------------------------------------------------------------
%	Glossaries
%-----------------------------------------------------------------------------

% Glossaries
%\usepackage[sanitize=none,acronym,toc]{glossaries}
\usepackage[acronym,toc]{glossaries}
%\usepackage[acronym,toc]{glossaries}
% Define a new glossary type
\newglossary[slg,toc]{symbols}{sym}{sbl}{Symbols List}
\newglossary[elg]{equations}{eqn}{eql}{Equations}
\makeglossaries

% ---------------------------------------------------------------------------- %
% Dicionario de Termos:
\input{d01-terms}
\input{d02-symbols}
\input{d03-acronyms}
\input{d04-equations}
% ---------------------------------------------------------------------------- %

% ---------------------------------------------------------------------------- %
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\erf}{erf}
% ---------------------------------------------------------------------------- %

%----------------------------------------------------------------------------------------
%	RUNNING HEAD AND CORRESPONDING AUTHOR
%----------------------------------------------------------------------------------------

% Author names in capital letters:
\authorrunninghead{PIRCHINER M. and ASSUMP\c{C}\~AO M.}

%------------------------------------------------

% Shorter version of title entered in capital letters:
\titlerunninghead{Smoothed seismicity Brazilian PSHA}

%------------------------------------------------

% Corresponding author mailing address and e-mail address:
\authoraddr{Corresponding author: 
Marlon Pirchiner, 
Department of Geophysics, IAG, USP, S\~ao Paulo, Brazil.
(marlon.pirchiner@iag.usp.br)}

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE
%----------------------------------------------------------------------------------------

\title{Smoothing seismicity techniques applied to
seismic source characterization and
probabilistic seismic hazard analysis in Brazil}

%----------------------------------------------------------------------------------------
%	AUTHORS AND AFFILIATIONS
%----------------------------------------------------------------------------------------

% Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote; matching \altaffiltext will appear at bottom of page.

\authors{Marlon Pirchiner\altaffilmark{1,2} 
		and Marcelo Assumpção\altaffilmark{1}
		}

\altaffiltext{1}{Department of Geophysics,
Institute of Astronomy, Geophysics and Atmospherical Sciences (IAG),
University of São Paulo (USP), São Paulo, Brazil.}

\altaffiltext{2}{Applied Math School (EMAp), Getúlio Vargas Foundation (FGV), Rio de Janeiro, Brazil.}

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

% Do NOT include any \begin...\end commands within the body of the abstract.

\begin{abstract}
Intraplate seismicity is a hard problem to solve in tectonics. In Brazil, as other intraplate tectonic settings, mainly during the lack of observed seismicity, the seismic hazard assessment are also difficult and the methods applied are often close correlated to the geometries defined by the experts with valuable criteria.

Three smoothed seismicity techniques are reviewed and their application on construct models of Brazilian seismicity and its applications to long-term seismic hazard assessment in Brazil are also discussed. All these smoothing methods are based on a \gls{kde}. The main difference between them is the bandwidth selection process.

The method proposed by \citet{frankel_1995} uses a \emph{fixed} bandwidth in space. \citet{woo_1996} suggests a bandwidth as function of the \emph{magnitude}. \citet{helmstetter_2012} propose a \emph{locally-adaptive} kernel bandwidths in space and time, chosen for each event as background (time-independent) model on their forecast experiments.

All these methods compute the seismicity rate surface in a binned space, time and magnitude, which able us to use this occurrence rate to define a grid of seismogenic point sources as the input source model to the hazard computation.

%The last two methods was implemented and are available as a open source software on the Hazard Modeling Toolkit (HMTK) from Global Earthquake Model (GEM) Foundation.

In addition, we used the information score and the error diagrams \citep{kagan_2009} to evaluate the models, optimize the smoothing distance on the \citet{frankel_1995} method, and finally give a weighted source models logic-tree to hazard computation.

All the hazard computations was performed on \gls{oq} \citep{pagani_2014} open source software from \acrfull{gem} Foundation. No focal mechanism information has been included or considered on these smoothed seismicity models. All hazard computations has been done using a single \gls{GMPE}. A more detailed study on \glspl{GMPE} selection should be developed further and it is beyond the scope of this work.

The results shown that all methods clearly most of the well known seismogenic regions and could be used in the Brazilian seismicity modeling. All methods improve the spatial resolution of the GSHAP hazard model, the only previously available for the region. These results should be considered on future reviews of Brazilian design building codes and procedures.

\end{abstract}

%------------------------------------------------------------------------------
%	ARTICLE CONTENT
%------------------------------------------------------------------------------
% The body of the article must start with a \begin{article} command
% \end{article} must follow the references section, before the figures and tables.

\begin{article}

\section{Introduction}

\subsection{Intraplate seismicity and previous studies in Brazil}

Brazil occupies a large stable continental region with seismicity rate lower than Central and Eastern North America, Australia, India or Northern Europe \citep{giardini_1999,johnston_1996a}. Spanning most of the central region of the South American plate, Brazil has earthquakes above magnitude 6 $mb$ with a return period of 50 years or so. Historical records are very short with almost no data before the XX century. Historical evidence of magnitude a 7 has been reported in the Amazon \citep{berrocal_1984,veloso_2014}. However, as in other intraplate settings, seismic hazard is not negligible due to the low attenuation in the cold and thick lithosphere (magnitude 5 $mb$ can be felt up to ~500 km away \citep{assumpcao_et_al_2014}) and special seismic hazard studies have been required for critical facilities such as nuclear power plants \citep{almeida_etal_2013}.

Few studies related to seismic hazard, such as \citet{berrocal_1996,santos_2010} for SE and NE Brazil, have been carried out so far and the first Brazilian seismic building code \citep{nbr_15421_2006,santos_lima_2008} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}. Brazilian seismicity is concentrated in several seismic zones \citep{berrocal_1984, assumpcao_et_al_2014}. Except for a peak of activity in Northeast Brazil, both the GSHAP model and the \citet{nbr_15421_2006} zoning (even when the Brazilian recommended levels are in line with those of other nations with higher occurrence of earthquakes \citep{santos_2013}) have not taken into proper account some of the other highly active seismic zones, such as the Porto dos Gaúchos, Mato Grosso state, with recurrent activity since the occurrence of the largest recorded Brazilian event with magnitude 6.2 $mb$ in 1955 \citep{barros_2009}, or the Central Amazon region where concentration of events in the last 50 years is consistent with the location of the possible historical earthquake of magnitude 7 \citep{veloso_2014}. For this reason, it is important to review and update the hazard estimates for Brazil, especially since the number of Brazilian earthquake records has been increased in the last three decades. %the increasing number of Brazilian earthquake records in the last three decades. 
%In addition, the results presented here are part of the \acrfull{gem} supported SARA project (South America Risk Assessment).

Despite attempts to explain the origin of some of the seismic zones, such as stress concentrations from lithospheric thinning \citep{assumpcao_2004, assumpcao_et_al_2014, azevedo_2015} or flexural effects \citep{zoback_1996, assumpcao_sacek_2013}, most of them have ill-defined boundaries. Attempts to produce seismic hazard maps \citep{dourado_2014} based on seismogenic zones are heavily dependent on subjective expert opinion, even more so in low seismic areas such as Brazil. Like a contribution to some new seismic hazard model for Brazil, here we briefly review recent methods of smoothing seismicity rates with application to Brazilian earthquakes.

%Despite some studies of seismicity rates and attempts at defining the main seismic zones \citep{berrocal_1984, berrocal_1996, assumpcao_et_al_2014}, Brazil has no seismic hazard map at national scale.

%The Brazilian seismic hazard building code \citep{nbr_15421_2006} was based on the GSHAP \citep{giardini_1999} results for South America \citep{tanner_shepherd_1997, shedlock_tanner_1999}, which is the last published model for the whole region. However, some significant seismic areas (Porto dos Gaúchos \citep{barros_2009}), are not represented in the GSHAP-based building codes.

%A significant number of earthquakes has been recorded in the last decades \citep{bsb_2014} allowing a better empirical and updated definition of the seismicity and hazard levels.

% definition of seismic zones. For this reason an updated hazard map is needed.

\subsection{Smoothed seismicity models}

Although the standard \citet{cornell_1968}--\citet{mcguire_1976} methodology often used to assess seismic hazard makes use of expert opinions to predefined seismogenic zones, one goal of this paper is to review a few smoothed seismicity methods \citep{riznichenko_1959} based on \gls{kde} which results in a grid or a smoothed seismicity surface without define any seismic zone polygon (also known as \textit{zoneless} approaches).

The (smoothed) seismicity \gls{kde} techniques allow us to define another grid of seismic point sources given an earthquake catalog. The seismicity rate (\emph{a}-value) associated to each source is defined by the smoothed seismicity methods.

To perform the \gls{kde}, both its shape and bandwidth needs to be defined. Most kernels are Gaussian, power-laws or one of their variants like Fisher kernels. The main distinction between the methods described here is the bandwidth selection process.

\citet{frankel_1995} smoothing method applies a Gaussian \citet{nadaraya_1964}-\citet{watson_1964} kernel smooth technique with a \emph{fixed smoothing-distance} bandwidth to estimate the seismicity rate by smoothing the 2D histogram from earthquakes counting. This algorithm is already available on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1}. 
%\citet{zechar_jordan_2010} integrate the contribution of each earthquake kernel on given cell grid and then evaluate each cell.
%The actual version of the algorithm is already implemented on the \gls{hmtk} \citep{weatherill_2012, weatherill_2014_1} with a simple variant of Frankel's method discussed by \citet{zechar_jordan_2010} which integrates the contribution of each earthquake kernel on given cell grid and then evaluate each cell instead smooth the counts \citep{frankel_1995}.

\citet{woo_1996} proposed a \emph{magnitude-dependent} kernel bandwidth function based on the regression for nearest neighbor mean distance between earthquakes binned by magnitude.

\citet{helmstetter_2012} proposed a \emph{locally-adaptive} bandwidth on both space and time dimensions for the background (time-independent) seismicity rate on their long-term forecasts experiments.

Is possible to use the long-term background seismic rate computed for forecast by its time-independence, the same \gls{psha}, assumption, even we preserve their different purposes \citep{marzocchi_2011}, and to apply this background seismicity models on seismogenic sources characterization \citep{weatherill_pagani_2014}.

\subsection{Information scores and \glspl{ed}}

From these three smoothing methods, only \cite{helmstetter_2012}, since its formalization, cross-validate their model parameters. All the other methods are concept to use the whole catalog and some fixed parameter (e.g. Frankel's smoothing distance).

To compare the performance of these three models, the information scores and error diagrams (\glspl{ed}) \citep{molchan_1992, kagan_2007, kagan_2009} were computed as alternative to the tests suggested to the \gls{relm} forecasts experiments \citep{schorlemmer_2007}.

On the evaluation phase all models learned from earthquakes into 1970-2004 observation time and were tested against the target realizations occurred between 2004 and 2014-01-01.

Scores from some Frankel-based models (from different smoothing distances) was computed and the best scores gave us an reasonable choice for this parameter.

HYBRID DECISION APPROACH BIRD/JORDAN[2013?]

\subsection{Hazard computation}

To perform the hazard computation, the \acrfull{oq} suite \citep{pagani_2014} was chosen. The modeling tool was the \gls{hmtk} also open available by \gls{gem} scientific board. The Frankel method was already implemented on the toolkit and two others was rewritten in the context of this paper based on sources given by the authors and their papers. % and should be opened on a public repository.

The other seismic source parameters, like \emph{b}-values, minimum and maximum magnitudes for the frequency-magnitude distributions, strike, dip and rate probabilities to the rupture planes on each source, its top and bottom limits on the subsurface, magnitude-area scale relationships and also the depth distributions must be assigned in other ways to proceed the hazard computation and take advantages of the latest generation \glspl{gmpe}.

In this work, to keep our focus on compute the hazard given a grid of punctual seismic sources computed from smoothed seismicity models, we compute the hazard probabilities using a single \gls{gmpe} [EXPOSE WHICH ONE] for a stable continental crust tectonic environment.

The earthquake catalog data comes from the Brazilian seismological research authorities \citep{bsb_2014}.

\subsection{Specific purposes}

The specific purposes of this paper are: 
(i) to present the \citet{bsb_2014} earthquake catalog,
(ii) review three distinct smoothing (zoneless) methods to spatially model the seismic rate,
(iii) evaluate each model by its information scores and error diagrams,
(iv) add into and share through \gls{hmtk} two more smooth methods to be used by the hazard modeling community,
(iii) use the open source \gls{gem}-\glsdesc{oq} tools to perform hazard calculations and
(v) update the hazard studies at national scale by proposing a simple hazard map based on these smoothed seismicity models.

Is not on our purposes at this time: (i) make an exhaustive evaluation of completeness magnitude in dependence of space and time, (ii) go deep on the \gls{gmpe} selection methodology and neither (iii) create the definitive Brazilian hazard map which we handle as a continuous process.

To go there, the Brazilian seismicity data will be presented and some preprocessing steps described. Next we will present the theory behind these smoothed seismicity methodologies. Its results will discussed in terms of its information scores. Then we will present the hazard computation formulating a the logic-tree as input for these source models. All this should allow us to take some lessons, trying to conclude or even suggest something which could still be developed in the future.

%------------------------------------------------
%
% seismicity data
%
%------------------------------------------------

\section{Brazilian seismicity data}

Data on crustal intraplate seismicity in Brazil is currently the Brazilian Seismic Bulletin maintained mainly by the Seismological Center of the University of São Paulo. This catalog is a joint compilation from different sources \citep{bsb_2014}.

\subsection{Epicenters and magnitudes}

For events up to 1981, epicenters in the catalog corresponds to \citet{berrocal_1984} which is a comprehensive compilation of historic and instrumental data with the participation from Universidade de São Paulo (USP), Universidade de Brasilia (UnB), Universidade Federal do Rio Grande do Norte (UFRN), Observatório Nacional (ON) and Instituto de Pesquisas Tecnológicas do Estado de São Paulo (IPT). From 1982 to 1995 the information comes from the annual bulletins published by Revista Brasileira de Geofísica. Since 1995, the Bulletin is maintained mainly by USP under the same cooperation network in addition to Universidade Estadual Paulista (UNESP) to compute the epicenters. Earlier versions of this catalog were used to compose the well-known \citet{ceresis_1985, ceresis_1995} catalogs.

Today the catalog is distributed in two ways. One, which could be called \emph{raw}, contains all compiled information, such as: events outside the country which were felt in Brazil, errors in previous versions and also non-seismic events (known quarry blasts, sonic boom, etc). The other way could be called \emph{clean}, where all non-seismic and low quality events were removed. This last one was used in this paper. There is 1941 earthquakes recorded from 1724 to the end of 2013.

This catalog only cover crustal events. Andean subduction earthquakes deeper than 50 km in the Brazil-Peru border was discarded as well earthquakes from the Mid-Atlantic ridge. Epicenters into the catalog are shown in figure \ref{fig_seismicity}.

Most magnitudes were computed as $mb$ type for teleseismic distances or the equivalent $mR$ \citep{assumpcao_1983} regional magnitude for short (300 to 1200 km) distances. For historical events magnitudes were estimated from macroseismic data (\glsdesc{sym:Af}, \gls{sym:Af} and \glsdesc{sym:I0}, \gls{sym:I0}) \citep{berrocal_1984}.

Since most \glspl{gmpe} are based on $M_W$ moment magnitude scale, in this present study, we converted all $mb$ magnitudes to $M_W$ values following the guidelines on table \ref{tab:guidelines}. REFERENCIA PARA STEPHANE/ASSUMPCAO WORK IN PROGRESS. (Johnston)

Other proxies for intraplate regions like \citet{scordilis_2006} was discarded by its magnitude range definition, over than 6, whilst most magnitudes in \citet{bsb_2014} catalog had values lower than this.

\subsection{Catalog checking}

Figure \ref{fig_overview} shows the distribution of depths, day-of-the-week and origin hour \citep{gulia_2012}. Most events have no depth determination and the well determined depths (from $pP$ phases or local networks) are mostly less than 10 km.

Figures \ref{fig_overview}b and \ref{fig_overview}c show an almost uniform distribution of day-of-the-week, and a slightly trend to record earthquakes during the night hours. This could suggest some influence of daily noise level on the station records.

The annual number of earthquakes since 1900 (figure \ref{fig_records}) shows a steady increase since 1960 and two peaks in 1986 and 1998 derived from two large seismic sequences: João Câmara \citep{takeya_1989} and Porto dos Gaúchos \citep{barros_2009}. The increasing number of records after 1960 is due to the operation of global, regional and local stations.

\subsection{Declustering procedures}

To preserve the Poisson assumption about the independence between distinct events, several decluster algorithms were tried and the results are shown in the figure \ref{fig_decluster}a. The window-method \citep{gardner_1974} was tested with three different windows: \citet{gardner_1974}, \citet{uhrhammer_1986} and Gr\"uenthal \citep{van_stiphout_2012}.
In addition the algorithm (AFTERAN) proposed by \citet{musson_1999} was also tested using Gr\"uenthal distance window. All this decluster methods are available in the \gls{hmtk}.

In the case of Brazilian catalog the methods did not present large differences at the final number of earthquakes (fig. \ref{fig_decluster}a) and the Gardner-Knopoff/Uhrhammer combination was chosen to preserve the maximum number of events.

The map on figure \ref{fig_decluster}b locates all clusters in the catalog. It is an \emph{a-priori} evidence of regions in Brazil where earthquake sequences tend to occur more frequently.

\subsection{Catalog completeness}

The completeness of Brazilian seismicity data has evident space and time dependence.

The number and quality of stations increased over time. The population density, previously concentrated along the coast increased also on other regions changing the historical instruments coverage. Brazil just recently installed its permanent seismographic network \citep{pirchiner_2011}.
%Trace back this history is a huge work with the current available station metadata, mainly from the analog stations.

There are many methodologies to assessment of the spatial and temporal completeness \citep{mignan_2014, vorobieva_2013, nasir_2013, mignan_2013, mignan_2012,  mignan_woessner_2012, mignan_2011, woessner_2005, stucchi_2004, cao_gao_2002, wiemer_wyss_2000, ogata_1993, stepp_1972}, or explaining how to compute the real \gls{pde} from the waveform records \citep{schorlemmer_2008}, but the time and efforts to proceed this exhaustive study of the \glsdesc{sym:Mc}, \gls{sym:Mc}, despite its relevance, is beyond the scope of this work. The waveforms from earliest years was not in digital form. The histories of most regional (temporary) stations are not easy available. In addition, the low number of earthquakes in the catalog also difficult some high resolution statistical methods. All of this hinders the direct application of most that methodologies. For this reason, the spatial completeness was not considered in this present work.

On the temporal dimension, for the whole country, two simple \gls{sym:Mc} methods were applied: the more subjective cumulative rate criteria or \gls{tcef} \citep{nasir_2013} which uses historical information about the records evolution and by the \citet{stepp_1972} method, an analytical approach, which uses the same kind information.

For the Stepp algorithm we use 0.5 and 2 as magnitude and time bins. The plot can be seen on the figure \ref{fig_completeness}a. The Stepp plot shows how each complete interval was set up by the point where, going back on time for each magnitude bin  --- increasing the number of samples, $T$ times 1 year samples --- the earthquakes by year ratio, $\lambda$ (expected to be time-independent, and always the same value), the standard error instead decrease as $1/\sqrt{T}$ samples, diverges. We also enforce the condition whilst greater magnitudes should have bigger observation times.

Figure \ref{fig_completeness}b shows the magnitude by time distribution in a scatter plot as well the cumulative number of events which is important to perform the \gls{tcef}. A linear behavior of cumulative rate on the recent years, gives an idea about the constant record trend. For magnitudes greater than 3, this threshold occurs about 1970. For the magnitudes greater than 4 and 5, these limits could be extrapolated to 1950 and 1940, respective and roughly. Figure \ref{fig_completeness}b also show the completeness tables from the \citet{stepp_1972} test and from the literature \citep{assumpcao_et_al_2014}. Tables \ref{tab:completeness_stepp} and \ref{tab:completeness} show each one of these completeness tables.

\subsection{Frequency-magnitude distribution}

Figure \ref{fig_occurrence} presents the catalog \gls{mfd}. For the minimum magnitude threshold was 2.9. The blue exes and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity. Stepp completeness (table \ref{tab:completeness_stepp}) and \citet{bsb_2014} declustered catalogue was used.

Simple \citet{gutenberg_richter_1944} incremental \glspl{mfd},
\begin{equation}
\log{N}(m|a,b) = a - b \times m,
\end{equation}
were fitted using a fixed $b$-value (b=1), showed as the blue continuous line (fig. \ref{fig_occurrence}), by the maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} methods (orange, $b=0.76\pm0.04$) are shown as continuous lines. The respective $a$-values was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.

From these distributions over the incremental catalog data, its possible to note just few records with magnitudes greater than 5, even considering different magnitude completeness on time, due probably the small observation time (few than 100 years) on Brazil stable crustal set.

Other important feature extracted from figure \ref{fig_occurrence}, is the low values of the $b$-values relative to the unity. Incremental b=1 \gls{mfd} underestimate the earthquake occurrences, specially under great magnitudes. From now on, we will assume the $b$-value of 0.9 as better representative for the whole region. On the future more investigations on the spatial $b$-value variations should be undertaken.

%------------------------------------------------
%
% smooothing methods
%
%------------------------------------------------
\section{Smoothing methods}
The smoothing methods considered in this article are now briefly explained and their results are discussed.

\subsection{Frankel: fixed bandwidth}

\citet{frankel_1995} smoothing seismicity proposal consisted originally in using a ``correlation distance'' $d_F$ as the \textit{fixed} kernel bandwidth and applying the Nadaraya-Watson \citep{nadaraya_1964, watson_1964} estimator to smooth a 2D seismicity histogram using a Gaussian kernel:
\begin{equation}
	\ensuremath{
		\tilde{n}_j = \frac{ \sum_{i} n_i \,e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}}
						   { \sum_{i}     e^{ - \left(\frac{\gls{sym:dij}}{\gls{sym:dF}}\right)^2}},
	}
	\label{eq:ni}
\end{equation}
where $\tilde{n}_j$ is the smoothed seismicity (number of earthquakes with magnitude $m$ above the minimum magnitude \gls{sym:Md} in the catalog) on cell $j$. $n_i$ is the earthquake counting in each other cell $i$ and \gls{sym:dij} is \glsdesc{sym:dij}.

\subsection{Woo: magnitude-dependent bandwidth}

\citet{woo_1996} suggested to evaluate the contribution of each earthquake $i$, located at $\boldsymbol{r}_i$, to the seismicity rate $R$ (number of earthquakes per year and unity area) on the cell centered in $\boldsymbol{r}$, depending on the magnitude $m$:
\begin{equation}
	\ensuremath{
		\gls{sym:Rrm} = \sum_{i=1}^{N} \frac{ K(\gls{sym:r} - \gls{sym:ri}, m)}
											{T({\gls{sym:ri})}},
	}
	\label{eq:Rrm}
\end{equation}
where $N$ is the number of earthquakes $i$ in the magnitude bin $m \pm \mathrm{d}m$,
and $T(\gls{sym:ri})$ is the completeness time of magnitude $m$ observed on \gls{sym:ri}.

Any kernel could be applied on that definition. I used used the \citet{kagan_knopoff_1980} kernel for infinite spatial domain as described by \citet{woo_1996}:
\begin{equation}
	\ensuremath{
		K_{KK}(\gls{sym:r}, m \arrowvert \gls{sym:aW}) =  \frac{  \gls{sym:aW}  -1}{\pi\gls{sym:hm}^2}
							\left[ 1 + \frac{\gls{sym:r}^2}{\gls{sym:hm}^2} \right]^{-\gls{sym:aW}},
	}
	\label{eq:k_kj}
\end{equation}
where \gls{sym:aW} is \glsdesc{sym:aW} \citep{verejones_1992}.

VERIFICAR SE É O CASO DE APRESENTAR OUTRAS FORMAS DE KERNEL. DAR PREFERENCIAS AOS REALMENTE UTILIZADOS.

To compute the magnitude-dependent bandwidth function \gls{sym:hm}, \citet{woo_1996} suggested the follow relation
\begin{equation}
	\ensuremath{
		h(m\arrowvert \gls{sym:a0}, \gls{sym:a1}) = \gls{sym:a0}e^{\gls{sym:a1}m},
	}
	\label{eq:hm}
\end{equation}
where $a_0$ and $a_1$ are computed by the regression of the mean nearest distance $h$ from each magnitude bin $m \pm \mathrm{d}m$, as illustrated on figure \ref{fig_woo_bandwidth}.

Just to compare, in figure \ref{fig_woo_bandwidth} it is shown the functions computed by \citet{beauval_2003} for Norway and Spain. The angular coefficient ($a_1$) is in the same range of the others but the general mean distance between events into the same magnitude bin is higher than other countries. It is in some way expected. Brazil is quite large than mentioned countries.

\subsection{Helmstetter and Werner: locally-adaptive space and time bandwidths}

Even work on a forecast perspective, the background seismicity rate for a long-term time-independent forecast could be used to characterize a diffused seismicity under the common assumption that this background seismic rate is invariant over time.

\citet{helmstetter_2012} proposes a seismicity model space and time dependent using kernels for space and time independently:
\begin{equation}
	\ensuremath{\gls{sym:R} = \sum_{i=1}^{N}{ \frac{1}{h_i\,{d_i}^2} \gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms01}
\end{equation}
where \gls{sym:R} is the \glsdesc{sym:R},
	  $K_t$ is the \glsdesc{sym:Kt},
	  $K_r$ is the \glsdesc{sym:Kr}.

As they are interested in forecast, just past time $t_i < t$ need to be considered. Also the observation completeness is taken in account by a set of weights $w$ in this follow way:
\begin{equation}
\ensuremath{\gls{sym:R} = \gls{sym:Rmin} + \sum_{t_i < t}{
	\frac{2\,w(\boldsymbol{r}_i,t_i)}{h_i\,{d_i}^2}
			\gls{sym:Kt}\gls{sym:Kr} }},
	\label{eq:helms02}
\end{equation}
where \gls{sym:Rmin} is the \glsdesc{sym:Rmin}, positive, allowing earthquakes to occur where their never occurred yet.

The weights $w(\boldsymbol{r}_i,t_i)$, computed for each earthquake $i$, are the Gutemberg-Richter \emph{a}-value projection. This is the expression used to compute it:
\begin{equation}
	\ensuremath{ w(\boldsymbol{r},t) = 10^{ \gls{sym:b}(\boldsymbol{r},t) \left[ \gls{sym:Mc_rt} - \gls{sym:Md}
	\right] } },
	\label{eq:helms_wi}
\end{equation}
where \gls{sym:wi} is the \glsdesc{sym:wi} factor computed on location $\boldsymbol{r}$ on instant $t$,
	  \gls{sym:b}$(\boldsymbol{r},t)$ is the space and time dependent \glsdesc{sym:b},
	  \gls{sym:Mc_rt} is the \glsdesc{sym:Mc_rt},
	  \gls{sym:Md} is the \glsdesc{sym:Md}.

These weights increase the seismicity contribution from earthquakes occurred where and when the completeness magnitude $M_c$ was greather than minimum catalog magnitude $M_d$. They also could easily take into account the space and time completenes and \emph{b}-value fluctuations.

\subsubsection{Local bandwidth computation}

The method implemented to compute the space and time bandwidths for each earthquake was the \gls{cnn} \citep{helmstetter_2012} expressed as:
\begin{equation}
	\ensuremath{
		h_i, d_i = \argmin_{\substack{h_i \ge \gls{sym:hk} \\
						              d_i \ge \gls{sym:dk}}
				           }
		\left[ s \left(h_i,d_i
			 		  \arrowvert
					  \gls{sym:k_cnn},\gls{sym:a_cnn}
			     \right)
			   := h_i + \gls{sym:a_cnn}d_i
	    \right],
	}
	\label{eq:helms_cnn}
\end{equation}
where \gls{sym:k_cnn} is the \glsdesc{sym:k_cnn},
	 \gls{sym:a_cnn} is a \glsdesc{sym:a_cnn},
	 \gls{sym:dk} is the \glsdesc{sym:dk} and
	 \gls{sym:hk} is the \glsdesc{sym:hk}.

The bandwidths are defined locally by this simple optimization process. It could be small on high earthquake density regions and higher where earthquakes are rarelly or regions with information lack.

\subsubsection{Stationary seismic rate}
After compute the model parameters it is completely defined, and the time-independent or stationary seismic rate $\bar{R}$ on each location $\boldsymbol{r}_0$ can be computed on the way proposed by \citet{helmstetter_2012} taking the median value for some location $\boldsymbol{r}_0$ over all considered time window:
\begin{equation}
	\ensuremath{
		\bar{R}(\boldsymbol{r}_0) = \text{Median}\left[R(\boldsymbol{r}_0, t)\right].
	}
	\label{eq:stationary_rate}
\end{equation}

The median should avoid the seismicity rate fluctuations derived by fore and after-shocks and consequently the decluster procedures. This is one of the most important achievement of this methodology.

\subsubsection{Maximum likelihood optimization}

The model is completely defined by \gls{sym:k_cnn}, \gls{sym:a_cnn} and \gls{sym:Rmin}. To optimize these parameters the catalog is divided in two parts: one for \emph{learning} about the parameters and other to test them. The \emph{testing} catalog will able us to check the model prediction performance derived from its chosen parameters. The best model parameters will maximize the prediction model capacity likely the target catalog.

If the earthquake occurrence could be modeled by a Poisson process with rate $N_p$, then the probability to observe exactly $n$ events on the considered time frame is
\begin{equation}
	\ensuremath{
		\gls{sym:pNn} = \frac{{N_p}^n e^{-N_p}}{n!}.
	}
	\label{eq:poisson_probability}
\end{equation}

Then, over all cells, the log-likelihood, to be maximized, between model prediction (from the learning catalog) and the observed earthquakes (on the testing catalog) is written as
\begin{equation}
	\ensuremath{
		\gls{sym:L} = \sum_{i_x=1}^{N_x}\sum_{i_y=1}^{N_y}\log p\left[  \gls{sym:Np}, \gls{sym:nxy}  \right]
	}
	\label{eq:loglik}
\end{equation}
where \gls{sym:Np} is the \glsdesc{sym:Np} and
	  \gls{sym:nxy} is the \glsdesc{sym:nxy}.

The model parameters $R_{min}$, $a_{cnn}$ and $k_{cnn}$ should be optimized by the maximization of log-likelihood $L$.

%------------------------------------------------

\section{Results}

.

\subsection{Simulations}

\subsubsection{Simulation 1}

.

%------------------------------------------------

\section{Discussion}

.

%----------------------------------------------------------------------------------------
%	APPENDICES (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here

% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing  \appendix

% \section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title

\appendix

\section{Appendix Title}

.

%----------------------------------------------------------------------------------------
%	GLOSSARY OR NOTATION (OPTIONAL)
%----------------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%------------------------------------------------------------------------------
\begin{acknowledgments}
The author thanks specially to 
%Marcelo Assumpção for the support and critical discussion as well 
all IAG-USP seismological team. It is also grateful to Stéphane Drouet and Vincent Guigues for all helpful discussions.
\end{acknowledgments}

%------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%------------------------------------------------------------------------------

% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}

% Or,

% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.

% Follow these steps:
% 1. Run LaTeX on your LaTeX file.

% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.

% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.

% 4. Comment out the old \bibliographystyle and \bibliography commands.

% 5. Run LaTeX on your new file before submitting.

% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

% ---------------------------------------------------------------------------- % Bibliografy
% ----------------------------------------------------------------------------
\bibliographystyle{BibTeX/agufull08} 	% citação bibliográfica textual
\bibliography{bib/bibliografia}  		% associado ao arquivo:

% \begin{thebibliography}{}

% \providecommand{\natexlab}[1]{#1}
% \expandafter\ifx\csname urlstyle\endcsname\relax
%   \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
%   \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
%   \urlstyle{rm}\Url}\fi

% \bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
% Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
%   logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
%   Comp.}, \textit{56}(193), 119--139.

% \bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
% Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
%   Scattering Theory}, John Wiley, New York.

% \bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
%   Wendland}}]{StephanHsiao}
% Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
%   problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
%   Appl. Math.}, \textit{34}(1), 1--19.

% \bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
% Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
%   components and line integrals of currents in modified edge representation,
%   \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

% \end{thebibliography}

% Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.

% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

\end{article}

%------------------------------------------------------------------------------
%	FIGURES AND TABLES
%------------------------------------------------------------------------------

%% Enter Figures and Tables here:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information should be placed in footnotes below the table.
%
%----------------
% EXAMPLE FIGURE
%
% \begin{figure}
% \noindent\includegraphics[width=20pc]{samplefigure.eps}
% \caption{Caption text here}
% \label{figure_label}
% \end{figure}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% \begin{figure}
% 	\begin{center}
% 	\begin{table}[H]
% 		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
% 		\footnotesize(a)
% 		&
% 		\footnotesize(b)
% 		\\
% 		\includegraphics[width=0.45\textwidth]{placeholder} &
% 		\includegraphics[width=0.45\textwidth]{placeholder}
% 		\end{tabular}
% 	\end{table}
% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
% 	\label{placeholder}
% 	\end{center}
% \end{figure}

% \begin{figure}
% 	\begin{center}
% 	\begin{table}[H]
% 		\begin{tabular}{ p{0.33\textwidth}
% 						 p{0.33\textwidth}
% 						 p{0.33\textwidth} }
% 		\footnotesize(a) &
% 		\footnotesize(b) &
% 		\footnotesize(c)
% 		\\
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_frankel} &
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_woo} &
% 		\includegraphics[width=0.33\textwidth]{z_img_pga_helmstetter}
% 		\end{tabular}
% 	\end{table}
% 	\caption{This tabled figure shows the first (a) image and the 2nd one (b). Even with the same bla ba.}
% 	\label{placeholder2}
% 	\end{center}
% \end{figure}

% See below for how to make sideways figures or tables.

\newpage

% =========
% intro
% =========

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_seismicity}
	\caption{Brazilian seismicity from 1767 to 2014-01-01. Source: \citep{bsb_2014}. The lack of information about the earthquakes' depths on this figure is due to almost none precise computation for them. When there is, the depths are around 10-15 km or shallower.}
	\label{fig_seismicity}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.20\textwidth}
						 p{0.40\textwidth}
						 p{0.40\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_depth}  &
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_weekday} &
		\includegraphics[height=5cm]{z_img_hmtk_bsb2014_11_hour}
		\end{tabular}
	\end{table}
	\caption{Catalog overview. The histogram (a) shows the depth distribution. Histograms (b) and (c) represent respectively the distributions of day-of-the-week and day hour which earthquakes occur. Dashed lines represent the mean value.}
	\label{fig_overview}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.99\textwidth]{z_img_hmtk_bsb2014_11_rate}
		\\
		\footnotesize(b)
		& \includegraphics[width=0.90\textwidth]{z_img_time_mag_count_br_1960}
		\end{tabular}
	\end{table}
	\caption{Earthquakes by year (a) and discriminated by magnitude (b). The minimum magnitude is 2.8. The time bin is 1 year for (a) and 2 years for (b). The magnitude bin on (b) is 0.5 unity. CHECK MAGNITUDE 6 MISSED EVENTS ON (b). }
	\label{fig_records}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.70\textwidth]{z_img_hmtk_bsb2014_11_decluster}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.60\textwidth]{z_img_hmtk_bsb2014_11_clusters}
		\end{tabular}
	\end{table}
	\caption{Declustering evaluation (a) using distinct algorithms and windows. There is no significant differences between final results. The map (b) of clusters from selected method shows places where earthquake sequences tends to occur.}
	\label{fig_decluster}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
        & \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_stepp}
		\\
		\footnotesize(b)
		& \includegraphics[width=0.99\textwidth]{z_img_completeness_temporal_scatter}
		\end{tabular}
	\end{table}
	\caption{Completeness evaluation. 
	The Stepp plot (a) was made using $\Delta_{mag}=0.5$ and $\Delta_{time}=2$ as magnitude and time bins. The magnitude-time scatter plot (b) shows in addition, three diferent temporal completess models and some cumulative rates to better perform a visual analysis.}
	\label{fig_completeness}
	\end{center}
\end{figure}

\begin{figure}
    \includegraphics[width=.99\linewidth]{z_img_bsb2014_recurrence}
    \caption{Annual earthquake recurrence. The blue x and red crosses are the incremental and cumulative recurrence distributions binned by 0.2 magnitude unity. Stepp completeness table and \citet{bsb_2014} declustered catalogue was used. Also a fixed b-value (b=1) incremental distribution is shown in blue continuous line, and the fitted by maximum-likelihood method (magenta, $b=0.91\pm0.07$), \citet{weichert_1980} (green, $b=0.88\pm0.04$) and \citet{kijko_2012} (orange, $0.76\pm0.04$) are shown as continuous lines. The respective $10^a$ was $3.7$, $3.8\pm0.2$, $3.5\pm0.2$, $3.2\pm0.1$.
    }
    \label{fig_occurrence}
\end{figure}

% =========
% Methods
% =========

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_woo_bandwidth}
	\caption{Magnitude dependence bandwidth.}
	\label{fig_woo_bandwidth}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_woo_kernels}
	\caption{\citet{kagan_knopoff_1980} kernel shapes for some magnitude values.}
	\label{fig_woo_kernels}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_hidi}
	\caption{Local bandwidth example.}
	\label{fig_helmstetter_hidi}
\end{figure}

% TODO add HI_DI histograms

\begin{figure}
	\begin{center}
	\begin{table}[H]
		%\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\begin{tabular}{ c c }
		\footnotesize(a)
		& \includegraphics[width=0.99\textwidth]{z_img_hi_histogram}
		\\
		\footnotesize(b)
        & \includegraphics[width=0.99\textwidth]{z_img_di_histogram}
		\end{tabular}
	\end{table}
	\caption{(a) $h_i$ and (b) $d_i$ distributions from \gls{bsb}.}
	\label{fig_hidi_histogram}
	\end{center}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_stationary_a}
	\caption{Stationary seismic rate}
	\label{fig_helmstetter_stationary_rate}
\end{figure}

\begin{figure}
	\includegraphics[width=0.99\linewidth]{z_img_helmstetter_catalogues}
	\caption{Learning and testing catalogs.}
	\label{fig_helmstetter_catalogues}
\end{figure}

% =========
% Smoothed Rates
% =========

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.25\textwidth}
						 p{0.25\textwidth}
						 p{0.25\textwidth}
						 p{0.25\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c) &
		\footnotesize(d)
		\\
		\includegraphics[height=4cm]{z_img_2dhist_assump_25}  &
		\includegraphics[height=4cm]{z_img_f_assump_50_25} &
		\includegraphics[height=4cm]{z_img_f_assump_150_25} &
		\includegraphics[height=4cm]{z_img_woo_assump_25} 
		\\
		\footnotesize(e) &
		\footnotesize(f) &
		\footnotesize(g) &
		\footnotesize(h)
		\\
		\includegraphics[height=4cm]{z_img_2dhist_stepp_25}  &
		\includegraphics[height=4cm]{z_img_f_stepp_50_25} &
		\includegraphics[height=4cm]{z_img_f_stepp_150_25} &
		\includegraphics[height=4cm]{z_img_woo_stepp_25} 
		\\
		\footnotesize(i) &
		\footnotesize(j) &
		\footnotesize(l) &
		\footnotesize(m)
		\\
		\includegraphics[height=4cm]{z_img_2dhist_assump_50}  &
		\includegraphics[height=4cm]{z_img_f_assump_50_50} &
		\includegraphics[height=4cm]{z_img_f_assump_150_50} &
		\includegraphics[height=4cm]{z_img_woo_assump_50} 
		\\
		\footnotesize(n) &
		\footnotesize(o) &
		\footnotesize(p) &
		\footnotesize(q)
		\\
		\includegraphics[height=4cm]{z_img_2dhist_stepp_50}  &
		\includegraphics[height=4cm]{z_img_f_stepp_50_50} &
		\includegraphics[height=4cm]{z_img_f_stepp_150_50} &
		\includegraphics[height=4cm]{z_img_woo_stepp_50}

		\end{tabular}
	\end{table}
	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
	\label{fig_rates2}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.33\textwidth}
						 p{0.34\textwidth}
						 p{0.33\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_a_frankel_br}  &
		\includegraphics[height=5cm]{z_img_a_woo} &
		\includegraphics[height=5cm]{z_img_a_helmstetter}
		\end{tabular}
	\end{table}
	\caption{Smoothed rates comparission: (a) shows \citet{frankel_1995} method smoothed rate results, (b) and (c) represent rates smoothed by \citet{woo_1996} and \citet{helmstetter_2012} methods respectly.}
	\label{fig_rates}
	\end{center}
\end{figure}

% =========
% Results
% =========

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.33\textwidth}
						 p{0.34\textwidth}
						 p{0.33\textwidth} }
		\footnotesize(a) &
		\footnotesize(b) &
		\footnotesize(c)
		\\
		\includegraphics[height=5cm]{z_img_pga_frankel}  &
		\includegraphics[height=5cm]{z_img_pga_woo} &
		\includegraphics[height=5cm]{z_img_pga_helmstetter}
		\end{tabular}
	\end{table}
	\caption{PGA (poe 10\%/50y) comparission for (a) \citet{frankel_1995}, (b) \citet{woo_1996} and (c) \citet{helmstetter_2012} methods.}
	\label{fig_pga}
	\end{center}
\end{figure}

\begin{figure}
	\begin{center}
	\begin{table}[H]
		\begin{tabular}{ p{0.5\textwidth} p{0.5\textwidth} }
		\footnotesize(a)
		&
		\footnotesize(b)
		\\
		\includegraphics[width=0.45\textwidth]{z_img_pga_gshap} &
		\includegraphics[width=0.45\textwidth]{z_img_pga_helmstetter}
		\end{tabular}
	\end{table}
	\caption{Previous available Brazillian PSHA from (a) GSHAP project and the PSHA proposed by this work (b). }
	\label{placeholder}
	\end{center}
\end{figure}

% TABLES

\begin{table}[H]
	\caption{BSB 2014.11 $M_W$ proxies \emph{ad-hoc} guidelines, sorted by priority. \gls{sym:Af} is the \glsdesc{sym:Af}, \gls{sym:I0} is the \glsdesc{sym:I0} in \gls{mmi} \citep{wood_1931}.}
	\begin{tabular}{l l l}
		\hline
		\textbf{magnitude source} & \textbf{rule} & \textbf{uncertainty}\\
		\hline
		$mb$ and $A_f$
		& $M_W(m, A_f) = 0.7 M_W(m) + 0.3 M_W(A_f)$
		& 0.33 \\
		% mb or equival & $$ M_W = 1.21m - 0.76 $$  & 0.3 \\
		% 		% magnitude from $A_f$  & $$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 8.4\times10^{-4}A_f^{1/2}& 0.4 \\
		$mb$ or $mR$
		&$ M_W(m) = 1.121 m - 0.76 $
		& 0.3 \\
		Felt area $A_f$
		&$ M_W(A_f) = 0.81 $$    + 0.639\log(A_f) + 0.00084\sqrt{A_f}$& 0.4 \\
		Maximum intensity $I_0$
		& $mb(I_0) = 1.21 + 0.45 I_0$ then $M_W(m)$
		& 0.6 \\
		\hline
	\end{tabular}
	\label{tab:guidelines}
\end{table}

\begin{table}[H]
	\caption{Temporal completeness. For each magnitude interval, the completeness was computed by the \citet{stepp_1972} method with time-step of 2 years and magnitude bin of 0.5 unity.}
	\begin{tabular}{c c }
		\hline
	\textbf{magnitude} & \textbf{year (Stepp)} \\
		\hline
		3.0 & 1970 \\
		% 3.5 &  	   \\
		4.0 & 1959 \\
		4.5 & 1951 \\
		% 5.0 &      \\
		% 5.5 &  	   \\
		6.0 & 1933 \\
		\hline
	\end{tabular}
	\label{tab:completeness_stepp}
\end{table}

\begin{table}[H]
	\caption{Temporal completeness. For each magnitude interval, the completeness comes from \citet{assumpcao_et_al_2014} converted to $M_W$ from $mb$ following table \ref{tab:guidelines}.}
	\begin{tabular}{c c}
		\hline
	\textbf{magnitude} & \textbf{year (Assumpção)} \\
		\hline
		3.1 & 1980 \\
		4.3 & 1968 \\
		4.8 & 1962 \\
		% 5.0 &      \\
		% 5.5 &  	   \\
		6.0 & 1940 \\
		\hline
	\end{tabular}
	\label{tab:completeness}
\end{table}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 4 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2 \]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with the first character following the equal sign on the first line.

% Use an \hspace{} command to insert horizontal space into your equation if necessary. Place an appropriate unit of measure between the curly braces, e.g. \hspace{1in}; you may have to experiment to achieve the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}